---
title: "Climate Change Drought Ratings"
author: "Hardy Griesbauer"
date: "09/10/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(forestDroughtTool)
library(tidyverse)

```

## Estimating future drought hazard under climate change scenarios
The Stand-Level Drought Tool estimates future drought by running the daily water balance model and adjusting daily cliamte data based on future normals.  This involves the following:

1. Use the BGC climate normal and future summaries produced by Will Mackenzie and Kiri Daust;
2. Determine the % difference between current (1961-1990 period) and future scenarios for monthly temperatures and precipitation; and
3. Apply those differences to daily climate data to estimate future daily data.

A few notes:
- Future scenarios in the BGC summaries are 4.5 and 8.5 scenarios.  We will also generate an 'average' scenario that averages the 4.5/8.5 data.

*Note: created an asmrFuture() function in forestDroughtTool package to handle all of this.*

### Step: Get normal (1961-1990) asmr for BGC units

```{r}
# What are the BGC units of interest?
load(here::here("dat","asmrNorth.rda"))
  
# BGC list
bgcList=asmrNorth$current$class$bgc
bgcList     

```

### Define a function to calculate future ASMR for each BGC using the futureASMR()

```{r}





```



